GRB Afterglows In The Multi-Messenger Era OR Electromagnetic Afterglows Of Structured Jets OR Structured Jets At All Angles

### 2019-05-16 structuredJets.tex
small updates.

\newcommand{\hubble}{{\em Hubble Space Telescope}}
\newcommand{\planck}{{\em Planck}}

@@

@@

@@

A generic energy profile $E(\theta)$ may have several more parameters, but one can still associate an on-axis energy $E_0 = E(0)$ and an effective core width $\thC \sim |E''(0)/E_0|^{-1/2}$ which should fully specify the leading order near on-axis behaviour.

@@

%%%%%%%
% METHODS
%%%%%%%

@@

To accommodate an initial structure profile $E(\theta)$ we consider the integrand of Equation \eqref{eq:flux} as a function of the polar angle $\theta$.  We assume each constant $\theta$ annulus evolves independently, as an equivalent top hat of initial width $\thC = \theta$.  This is a very good approximation when transverse velocities are low: when the blast wave is ultra-relativistic and has not begun to spread and when the blast wave is non-relativistic and the spreading has ceased \citep{van-Eerten:2010aa}.  Once jet spreading begins in earnest the errors are larger, and this approach can be best viewed as an interpolation between the correct ultra-relativistic and non-relativistic limits \citep{van-Eerten:2010aa}.

@@

Comparing to \boxfit{} tests the fidelity of some of the assumptions and approximations of the single shell approach used by \afterglowpy{}.  The simpler analytic version of the scheme on which \afterglowpy{} is based differed from \boxfit{} by factors from 2 to 8 \citep{vanEerten:2010aa}.  We find \afterglowpy{} performs much better, with a relative difference of less than 50\% for the bulk of the evolution.  The difference is worst, up to a factor of 3 relative to \boxfit{}, at the onset of jet spreading.

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{figs/lc_thV_model_multi.pdf}
	\caption{Aligned ($\thobs = \thC/2$) afterglow light curves for top-hat (blue), Gaussian (orange), and power law (green) jets. When viewed on-axis, jets containing angular structure are very difficult to distinguish from top-hats.  The transition phase marking the onset of the jet break begins at $\tbm\sim3$ days with a shallow break and ends sharply at $\tbp \sim 100$ days.  These particular light curves use $\thC = 0.1$ rad, $\thW = 0.4$ rad, $E_0 = 10^{52}$erg, $n_0=10^{-2}$cm$^{-3}$, $p=2.2$, $\epse = 10^{-1}$, $\epsB = 10^{-3}$, and $d_L=40$Mpc. \label{fig:onaxis}}
\end{figure*}

@@

\section{Refreshed Shocks --- Fast Tails and Spherical Cocoons}\label{sec:refreshedShocks}

@@

\section{Relation to Energy Injection Models}

@@

@@

>>>>

\newcommand{\hubble}{{\em Hubble Space Telescope}}
\newcommand{\planck}{{\em Planck}}
\newcommand{\beppoSAX}{{\em BeppoSAX}}

@@

% Present new closure relations for structured jets, compare with established relations for choked jet models...

@@

\begin{itemize}
\item \citet{Meszaros:1998aa} First discusses effect of angular structure in $E$ and $\Gamma$, predicts rising light curves when seen off axis, calculates closure relations in various cases.
\item \citet{Dalal:2002aa} Considers emission from boosted lightbulb, refers to $\thobs > \theta_j$ as ``off-axis,'' points out off-axis jet breaks will be delayed and dim, light curves converge at late time.
\item \citet{Moderski:2000aa} Relativistic effects (equal arrival time surface, etc) smear jet-break.
\item \citet{Rossi:2002aa} Considers jets with $E\propto \theta^{-2}$.  Verifies at early times line-of-sight material dominates over core and wings, light curves resemble on-axis top-hats, but jet break is due to observer angle.  Simple emission model.  Postulates variety of jet-breaks due to off-axis structured jets, instead of distribution of $\Eiso$, $\thC$.  $t_b \propto \thobs^{8/3}$.  
\item \citet{Granot:2002ab} Off-axis viewing of top hat jets, computed as boosted light bulb, analytic top-hat, and hydro sim. $\thobs < \theta_j$ look like on-axis top-hats, jet breaks increase with $\thobs$, evidence for $\thobs$-dependent slope before jet break in hydro models.
\item \citet{Lipunov:2001aa, Zhang:2002aa} universal structured jet, $\theta^{-2}$, may explain jet breaks, etc.
\item \citet{Panaitescu:2003aa} Off-axis viewing of power-law jets $E\propto \theta^{-q}$ with sharp cores.  Analytic blast waves, detailed treatment of relativistic emission and synchrotron spectrum. When viewing within the core, no large effect so long as $q \geq 2$.  Analytic formulae for aligned observers, numerical only for misaligned.  Core visible at $t_a \sim E_0^{1/3} \thobs^{8/3}$ in ISM.  Structure may only steepen jet breaks, proposes examining pre- and post- jet break slopes to determine whether structure is evident.
\item \citet{Kumar:2003aa} Numerical evolution of radially-integrated Gaussian, top-hat, and power-law jets/  ``Therefore, a simple model in which the energy per unit solid angle is taken to be time independent and each element of the jet behaves as if it were part of a spherical flow with the same can serve as a useful approximation for the jet dynamics, as long as the jet is sufficiently relativistic.''
\item \citet{Granot:2003aa} ``structured jet:'' jet with angular structure. Smooth power law in $\Theta = \sqrt{1 + \theta^2/\thC^2}$.  Analysis of power-law jet light curves from \citet{Kumar:2003aa}. Only considers USJ, top-hat, and permutations.  Find deviations from top-hat appearance small, USJ must have $E\approx \theta^{-2}$, $\Gamma\approx \theta^{0 < 1}$.
\item \citet{Salmonson:2003aa} Numerical and analytic study of top-hat and structured jets.  Computes doppler factor and angular size in asymptotic limits (no cooling frequency), finds far-off-axis closure relations.  Examines `bump' seen when initially strongly beamed emission becomes structured, then proceeds to jet break.
\item \citet{Nakar:2004aa} Distribution of luminosity and jet-breaks rules out USJ. 
\item \citet{Rossi:2004aa}. Light curves and polarization for top-hat, USJ, and Gaussian jet.  Gaussian jet shows slowly rising flux at sufficiently large viewing angles. Time dependence of polarization signal may distinguish between USJ and Gaussian.
\end{itemize}

The electromagnetic afterglow of \gwbns{} gave, quite literally, a new view on gamma-ray burst (GRB) afterglows.

A canonical afterglow was almost immediately ruled out \citep{Troja:2017aa}.

The structure of the blast wave giving rise to gamma-ray burst (GRB) afterglows has been a subject of interest since their discovery by \beppoSAX{} in GRB 970228 \citep{Wijers:1997aa}.

{\color{blue} LUIGI: ?was found since early afterglow observations with BeppoSAX, when the
presence of a structured outflow has been considered for explaining some of the
temporal breaks observed in the light curves of a number of GRBs (Panaitescu and
Kumar 2003, Apj 592: 390, Panaitescu 2005 MNRAS 362, 921). Candidates for
structured jet exhibit a variation of the decay index at the break smaller that that
expected for a top-hat spreading jet ($\Delta \alpha \approx 0.8$) and larger than
that induced by the crossing of the cooling frequency ($\Delta\alpha=0.25$). Energy
injection by a long lasting central engine can also induce temporal breaks, either
flattening or steepening. Other effects such as transition to non-relativistic
expansion, the transition from wind-like to ISM (Piro et al. 2005, Apj 623 314) or the
emergence of an Inverse compton component (Corsi \& Piro
2006A\&A...458..741C) can also produce breaks, in these case associated with a
flattening of the light curve.
More recenlty Ryan et al (2015) presented evidence of a structured jet in a sample
based on SWIFT observations. Considering the number of processes that can
produce breaks mentioned above, no undisputable observational evidence of a
structured jet became available till the observations of GW170817. Real jets}

Gamma-ray burst (GRB) afterglows, since their discovery by \beppoSAX{} in GRB 970228, have been a potent tool for studying GRBs and their progenitors.

@@

A generic energy profile $E(\theta)$ may have several more parameters, but one can still associate with it an on-axis energy $E_0 = E(0)$ and an effective core width $\thC \sim |E''(0)/E_0|^{-1/2}$ which should fully specify the leading order near on-axis behaviour.

@@

@@

To accommodate an initial structure profile $E(\theta)$ we consider the integrand of Equation \eqref{eq:flux} as a function of the polar angle $\theta$.  We assume each constant-$\theta$ annulus evolves independently, as an equivalent top hat of initial width $\theta_j = \theta$.  This is a very good approximation when transverse velocities are low: when the blast wave is ultra-relativistic and has not begun to spread and when the blast wave is non-relativistic and the spreading has ceased \citep{van-Eerten:2010aa}.  Once jet spreading begins in earnest the errors are larger, and this approach can be best viewed as an interpolation between the correct ultra-relativistic and non-relativistic limits \citep{van-Eerten:2010aa}.

@@

Comparing to \boxfit{} tests the fidelity of some of the assumptions and approximations of the single shell approach used by \afterglowpy{}.  The simpler analytic version of the scheme on which \afterglowpy{} is based differed from \boxfit{} by factors from 2 to 8 \citep{van-Eerten:2010aa}.  We find \afterglowpy{} performs much better, with a relative difference of less than 50\% for the bulk of the evolution.  The difference is worst, up to a factor of 3 relative to \boxfit{}, at the onset of jet spreading.

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{figs/lc_thV_model_multi.pdf}
	\caption{X-ray afterglow light curves from structured jets at different viewing angles $\thobs$, calculated with \afterglowpy{}.  Each panel shows the 1 keV flux density $F_\nu$ and the corresponding temporal slope $\alpha = d\log F_\nu / d\log \tobs$.  Included models are the top-hat jet (dashed blue), Gaussian jet (solid green), $b=2$ power law (dot-dashed orange), and $b=6$ power law at $\thobs=0$ (upper left), $\thobs=2\thC=0.16$ rad (upper right), $\thobs=4\thC=0.32$ rad (lower left), and $\thobs=6\thC=0.48$ rad (lower right). These particular light curves all use $\thC = 0.08$ rad, $\thW = 0.24$ rad, $E_0 = 10^{53}$erg, $n_0=1$ cm$^{-3}$, $p=2.2$, $\epse = 10^{-1}$, $\epsB = 10^{-2}$, $d_L=10^{28}$cm and $z=0.5454$. \label{fig:thVmodel}}
\end{figure*}

@@

\subsection{Refreshed Shocks --- Fast Tails and Spherical Cocoons}\label{subsec:refreshedShocks}

@@

\subsection{Relation to Energy Injection Models}

@@

%%%%%%%%%%%%%%%
%
%.  DISCUSSION
%
%%%%%%%%%%%%%%%

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{figs/fit_E_theta.pdf}
	\caption{Inferred $E(\theta)$ for \gwbns{} assuming Gaussian (blue, solid line) or power law (orange, dashed line) structure models.  Lines show the median posterior value for $E(\theta)$ at each fixed value of $\theta$, shaded regions show the symmetric 68\% confidence interval (16\% and 84\% quantiles) at each $\theta$. \label{fig:fit_E_theta.pdf}}
\end{figure}
